The $\lambda ^? _{\rightarrow}$ syntax is essentially the same as 
the simply-typed $\lambda$-calculus with references, except for the 
addition of dynamic types ?.
\begin{alignat*}{3}
    &(\textit{expressions}) \qquad 
    & e ::= \: & x \; | \;  
          e \: e \; | \;
          \lambda x \! : \! T . \: e \; | \;
          k \\
    &&    & \texttt{succ} \: e \; | \;
          \texttt{pred} \: e \; | \;
          \texttt{iszero} \: e \; | \;
          \texttt{if} \: e \: \texttt{then} \: e \: \texttt{else} \: e \; | \; \\
    &&    & \texttt{ref} \: e \; | \;
          !e \; | \;
          e := e \\
    &(\textit{constants}) 
    & k ::= \: & n \; | \; \texttt{true} \; | \; \texttt{false} \\
    &(\textit{numbers}) 
    & n ::= \: & 0 \; | \; \texttt{succ} \: n \\
    &(\textit{types}) 
    & T ::= \: & \texttt{Bool} \; | \;
            \texttt{Nat} \; | \;
            T \rightarrow T \; | \;
            \texttt{Ref} \: T \; | \;
            ?     
\end{alignat*}

Expressions include variables, abstractions, and applications, as well as 
arithmethic, boolean, and conditional terms found in the untyped $\lambda$-calculus. 
Reference expressions consist of allocation, dereference, and assignment.
Types include the base types (\texttt{Bool} and \texttt{Nat}), function 
types, reference types, and finally dynamic types. 
Constants like \texttt{true} and \texttt{false} inherently  
have \texttt{Bool} type, whereas numerical values are of type \texttt{Nat}. 
Function parameters without type annotations automatically 
assume dynamic types. 

The type system of $\lambda ^? _{\rightarrow}$ is based on the 
consistency relation between known and unknown types, given in 
Figure 1 [8]. Note that the consistency relation is reflexive and 
symmetric, but not transitive. Given $S \sim \: ?$ and $? \: \sim T$, it 
is not always the case that $S \sim T$.  

\input{figures/Consistency}

Figure 2 presents the type system of $\lambda ^? _{\rightarrow}$. Rules 
{\scriptsize{[T-VAR]}}, {\scriptsize{[T-CONST]}}, {\scriptsize{[T-FUN]}}, 
and {\scriptsize{[T-REF]}}  
are straightforward. The rules for function applications handle either 
known or unknown function types. Given a specific function type, an 
application only accepts an argument type that is consistent to the 
parameter type. \input{figures/TypeSystem} 
Otherwise, the argument can have any type, and the return 
type of the application is unknown. Similarly, there are two assignment 
rules for whether the left-hand side is dynamically-typed. 
A dereference expression only has a specific type if its argument is a 
reference. We treat arithmetic 
and boolean expressions ($\texttt{succ} \: e$, $\texttt{pred} \: e$, $\texttt{iszero} \: e$) 
as constant functions that accept any type consistent to \texttt{Nat} and 
return either a \texttt{Nat} or a \texttt{Bool}. Conditional expressions are more 
complicated because we have to typecheck both the condition and the two 
branches. The condition type must be consistent to 
\texttt{Bool}. The types of two branches need to be consistent. If so, the 
resulting type can derive from either branch type.
