\documentclass[11pt]{article}
\usepackage{afterpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{caption}
\usepackage[margin=4cm]{geometry}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{setspace}
\usepackage{textcomp}
\usepackage[hyphens]{url}
\usepackage{xcolor}

\allowdisplaybreaks

\newcommand\blankpage{
    \null
    \thispagestyle{empty}
    \addtocounter{page}{-1}
    \newpage}
    
\makeatletter
\newenvironment{shiftedflalign}{
  \start@align\tw@\st@rredfalse\m@ne
  \hskip\parindent
}{
  \endalign
}
\newenvironment{shiftedflalign*}{
  \start@align\tw@\st@rredtrue\m@ne
  \hskip\parindent
}{
  \endalign
}
\makeatother

\lstset{basicstyle=\small\sffamily,
        breaklines=true,
        language=Haskell,
        columns=flexible,
        keepspaces=true,
        tabsize=4,
        commentstyle=\color{gray},
        morekeywords={throwError},
        literate= {<-}{$\leftarrow$}{1}{->}{$\rightarrow$}{1},
        deletekeywords={fst,snd}
      }

\newenvironment{bprooftree}
  {\leavevmode\hbox\bgroup}
  {\DisplayProof\egroup}

\newcommand{\highlight}[1]{
  \colorbox{gray!15}{$\displaystyle#1$}}

\captionsetup{labelfont=bf,font=bf,skip=0pt}

\title{\bf Interpretation of the Gradually-Typed Lambda Calculus}
\author{Tien Thuy Ho}

\begin{document}

% Title page
\input{TitlePage}

% Copyright
\input{Copyright}

% Table of contents
\setcounter{page}{3}
\tableofcontents
\listoffigures

\maketitle
\pagenumbering{arabic}

\begin{abstract}
Gradual typing offers both type safety and programming flexibility by integrating static and dynamic type checking 
together within the same language. The runtime semantics of a gradually-typed system involves the translation of the 
source language to an intermediate one with explicit casts. Space comsumption of gradual typing can be improved by using 
the coercion calculus to collapse a sequence of adjacent casts into a single cast [4]. This paper details an efficient 
interpretation of the gradually-typed lambda calculus based on coercions with eager error detection, as described formally 
by Herman et al. [4]. We also augment the system with UD blame assignment [8] to pinpoint the particular expressions 
responsible for coercion failures.
\end{abstract}

\section{Introduction} \input{Introduction}
\section{Gradually-Typed Lambda Calculus $\lambda ^? _{\rightarrow}$} \input{GTLC}
\section{The UD Blame Assignment} \input{BlameAssignment}
\section{The Eager Coercion Calculus} \input{Coercion}
\section{Interpreter for $\lambda ^? _{\rightarrow}$} \input{Interpreter}
\section{Related Work} \input{RelatedWork}
\section{Conclusion} \input{Conclusion}

\appendix
\input{Appendix}

\begin{thebibliography}{9}
    \bibitem{diehl} 
    S. Diehl:  Parsing, \\
    \url{http://dev.stephendiehl.com/fun/002\char`_parsers.html}

    \bibitem{parsec-tutorial} 
    Haskell: Parsing a Simple Imperative Language, \\
    \url{https://wiki.haskell.org/Parsing\char`_a\char`_simple\char`_imperative\char`_language}

    \bibitem{henglein} 
    F. Henglein.
    Dynamic typing: Syntax and proof theory. 
    \textit{Sci. Comput. Program.},
    22(3):197-230, 1994.

    \bibitem{flanagan} 
    D. Herman, A. Tomb, and C. Flanagan. 
    Space-efficient gradual typing. 
    In \textit{Trends in Functional Prog.}, (TFP),
    pages XXVIII, April 2007.

    \bibitem{hutton} 
    Graham Hutton. 
    \textit{Programming in Haskell}. 
    University Press, Cambridge, 2007.

    \bibitem{pierce} 
    Benjamin C. Pierce. 
    \textit{Types and Programming Languages}. 
    The MIT Press, Cambridge, 2002.

    \bibitem{garcia} 
    J. G. Siek and R. Garcia. 
    Interpretations of the gradually-typed lambda calculus. 
    In \textit{Scheme and Functional Programming Workshop}, 
    pages 68-80, September 2012.

    \bibitem{taha} 
    J. G. Siek, R. Garcia, and W. Taha. 
    Exploring the design space of higher-order casts. 
    In \textit{European Symposium on Programming}, ESOP,
    pages 365-376, January 2010.

    \bibitem{siek} 
    J. G. Siek and W. Taha. 
    Gradual typing for functional languages. 
    In \textit{Scheme and Functional Programming Workshop}, pages 81-92, September 2006.

    \bibitem{source} 
    Source code, \\
    \url{https://github.com/thuytien140894/GTLC}
\end{thebibliography}
\end{document}